\section{The CARM Model}

\subsection{Flat Object Composition}

Let $K$ be a finite set of keys and $V$ a set of values. 
A blueprint is defined as a finite partial mapping:

\[
B : K \rightharpoonup V
\]

Domphy models a UI entity $E$ as the result of a deterministic merge over compositional layers:

\[
E = P \oplus B \oplus I \oplus R
\]

where:

\begin{itemize}
    \item $P$ (Presets) denotes system-level default mappings,
    \item $B$ (Behaviors) denotes modular structural augmentations,
    \item $I$ (Instance) denotes local attribute definitions,
    \item $R$ (Reactive mappings) denotes property-level state bindings.
\end{itemize}

The operator $\oplus : (K \rightharpoonup V) \times (K \rightharpoonup V) \to (K \rightharpoonup V)$ 
is defined as a right-biased associative merge:

\[
(A \oplus B)(k) =
\begin{cases}
B(k) & \text{if } k \in \text{dom}(B) \\
A(k) & \text{otherwise}
\end{cases}
\]

This merge preserves a flat key–value structure; composition does not introduce additional structural nesting beyond the underlying DOM hierarchy. Structural composition is therefore resolved at definition time rather than at evaluation time.
\subsection{Atomic Reactive Mapping}

Let $S$ denote the set of reactive signals.
For each reactive attribute $a \in A$, Domphy establishes a persistent mapping:

\[
f_a : S_a \to V_a
\]

where $S_a \subseteq S$ represents the subset of signals upon which $a$ depends.

The collection of mappings
\[
\mathcal{F} = \{ f_a \mid a \in A \}
\]
forms a dependency graph $G = (S \cup A, E)$, 
constructed during a one-time reification phase.
\subsubsection{Computational Efficiency}

Let $\Delta S \subseteq S$ denote the set of updated signals.
Let $A_{\text{affected}}$ be the set of attributes dependent on $\Delta S$.

If the dependency graph $G$ is static after reification and 
the maximum out-degree of any signal node is bounded by a constant $c$,
then the update cost satisfies:

\[
T(\Delta S) = O(|A_{\text{affected}}|)
\]

Under the bounded-dependency assumption ($|A_{\text{affected}}| \le c$ per signal),
the update cost per signal is $O(1)$.

Unlike tree-based reconciliation strategies that may require traversal of
structural subtrees, CARM propagates updates directly along pre-established
signal-to-attribute edges.
