\begin{thebibliography}{30}

% --- Modular scales and spatial systems ---

\bibitem{bringhurst2012}
R.~Bringhurst.
\newblock \emph{The Elements of Typographic Style}, Version 4.0.
\newblock Hartley \& Marks, 2012.

\bibitem{brown2011}
T.~Brown.
\newblock More meaningful typography.
\newblock \emph{A List Apart}, May 2011.
\newblock Available at \url{https://alistapart.com/article/more-meaningful-typography/}.

\bibitem{curtis2016}
N.~Curtis.
\newblock Space in design systems.
\newblock \emph{EightShapes}, September 2016.
\newblock Available at \url{https://medium.com/eightshapes-llc/space-in-design-systems-188bcbae0d62}.

\bibitem{kutschmann2024}
A.~Kutschmann.
\newblock Design token-based UI architecture.
\newblock Published on \emph{martinfowler.com}, December 2024.
\newblock Available at \url{https://martinfowler.com/articles/design-token-based-ui-architecture.html}.

\bibitem{w3cdt2025}
Design Tokens Community Group.
\newblock \emph{Design Tokens Format Module 2025.10}.
\newblock W3C Community Group Report, 2025.
\newblock Available at \url{https://www.designtokens.org/tr/drafts/format/}.

\bibitem{dahl2017}
E.~Dahl.
\newblock Intro to the 8-point grid system.
\newblock \emph{Built to Adapt}, 2017.
\newblock Available at \url{https://medium.com/built-to-adapt/intro-to-the-8-point-grid-system-d2573cde8632}.

% --- Density ---

\bibitem{materialdensity}
Google.
\newblock Applying density.
\newblock \emph{Material Design 2 Guidelines}, 2019.
\newblock Available at \url{https://m2.material.io/design/layout/applying-density.html}.

\bibitem{kravets2020}
U.~Kravets.
\newblock Using Material density on the web.
\newblock \emph{Google Design}, 2020.
\newblock Available at \url{https://medium.com/google-design/using-material-density-on-the-web-59d85f1918f0}.

\bibitem{gabillon2013}
Y.~Gabillon, S.~Lepreux, and K.~M. de~Oliveira.
\newblock Towards ergonomic user interface composition: A study about
  information density criterion.
\newblock In \emph{Human-Computer Interaction (HCI 2013)}, LNCS 8004,
  pages 216--225. Springer, 2013.

\bibitem{tufte2001}
E.~R. Tufte.
\newblock \emph{The Visual Display of Quantitative Information}.
\newblock Graphics Press, 2nd edition, 2001.

% --- Visual consistency and perceptual grouping ---

\bibitem{wertheimer1923}
M.~Wertheimer.
\newblock Untersuchungen zur Lehre von der Gestalt, {II}.
\newblock \emph{Psychologische Forschung}, 4:301--350, 1923.
\newblock English translation in W.~Ellis, \emph{A Source Book of Gestalt
  Psychology}, Routledge, 1938.

\bibitem{wagemans2012}
J.~Wagemans, J.~H. Elder, M.~Kubovy, S.~E. Palmer, M.~A. Peterson, M.~Singh,
  and R.~von~der~Heydt.
\newblock A century of {G}estalt psychology in visual perception: {I}.
  {P}erceptual grouping and figure--ground organization.
\newblock \emph{Psychological Bulletin}, 138(6):1172--1217, 2012.

\bibitem{nielsen1994}
J.~Nielsen.
\newblock Enhancing the explanatory power of usability heuristics.
\newblock In \emph{CHI '94: Proceedings of the SIGCHI Conference on Human
  Factors in Computing Systems}, pages 152--158. ACM, 1994.

\bibitem{shneiderman2016}
B.~Shneiderman, C.~Plaisant, M.~Cohen, S.~Jacobs, and N.~Elmqvist.
\newblock \emph{Designing the User Interface: Strategies for Effective
  Human-Computer Interaction}.
\newblock Pearson, 6th edition, 2016.

% --- Typography and line height ---

\bibitem{wcag21text}
W3C.
\newblock \emph{Web Content Accessibility Guidelines (WCAG) 2.1}.
\newblock 2018.
\newblock Success Criterion 1.4.12: Text Spacing.
\newblock Available at \url{https://www.w3.org/TR/WCAG21/#text-spacing}.

\bibitem{rello2016}
L.~Rello, M.~Pielot, and M.-C. Marcos.
\newblock Make it big! {T}he effect of font size and line spacing on online
  readability.
\newblock In \emph{CHI '16: Proceedings of the 2016 CHI Conference on Human
  Factors in Computing Systems}, pages 3637--3648. ACM, 2016.

% --- Design systems (industry validation) ---

\bibitem{mui}
MUI.
\newblock \emph{Material UI}.
\newblock 2024.
\newblock Available at \url{https://mui.com/}.

\bibitem{antdesign}
Ant Group.
\newblock \emph{Ant Design}.
\newblock 2024.
\newblock Available at \url{https://ant.design/}.

\bibitem{chakra}
S.~Adebayo.
\newblock \emph{Chakra UI}.
\newblock 2024.
\newblock Available at \url{https://chakra-ui.com/}.

\bibitem{primer}
GitHub.
\newblock \emph{Primer Design System}.
\newblock 2024.
\newblock Available at \url{https://primer.style/}.

\bibitem{spectrum}
Adobe.
\newblock \emph{Spectrum Design System}.
\newblock 2024.
\newblock Available at \url{https://spectrum.adobe.com/}.

\bibitem{material3}
Google.
\newblock \emph{Material Design 3}.
\newblock 2023.
\newblock Available at \url{https://m3.material.io/}.

\bibitem{css-box}
W3C.
\newblock \emph{CSS Box Model Module Level 3}.
\newblock 2023.
\newblock Available at \url{https://www.w3.org/TR/css-box-3/}.

% --- Additional design systems (benchmark) ---

\bibitem{fluentui}
Microsoft.
\newblock \emph{Fluent UI}.
\newblock 2024.
\newblock Available at \url{https://fluent2.microsoft.design/}.

\bibitem{carbon}
IBM.
\newblock \emph{Carbon Design System}.
\newblock 2024.
\newblock Available at \url{https://carbondesignsystem.com/}.

\bibitem{atlassian}
Atlassian.
\newblock \emph{Atlassian Design System}.
\newblock 2024.
\newblock Available at \url{https://atlassian.design/}.

\bibitem{polaris}
Shopify.
\newblock \emph{Polaris Design System}.
\newblock 2024.
\newblock Available at \url{https://polaris.shopify.com/}.

\bibitem{slds}
Salesforce.
\newblock \emph{Lightning Design System}.
\newblock 2024.
\newblock Available at \url{https://www.lightningdesignsystem.com/}.

% --- Chromametry ---

\bibitem{chromametry}
H.~K. Nguyen.
\newblock Web-accessible color palette metrics.
\newblock 2025.
\newblock Available at \url{https://github.com/chromametry/chromametry}.

\end{thebibliography}
